---
title: "Interactive Census Data"
author: "Jake Hannan"
date: "September 21, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### About the Census API

The US Census Bureau publishes a wealth of aggregated metrics for several geographic boundaries (e.g., zip codes, census block groups, census tracts) - this information can be used by anyone. 

Additionally, in order to visualize this data, the US Census Bureau offers spatial files at many levels of granularity (see https://www.census.gov/programs-surveys/geography.html for more). The `tigris` package is what we'll start with to find some county level data, and also use `spdplyr` to do some basic filtering on the data frame.

```{r, echo=TRUE, message=FALSE, results='hide', warning=FALSE}
require(tigris)
require(spdplyr)
#get US Counties using `tigris`
us_counties <- counties(state = NULL, year = 2017)
#filter areas not interested via `spdplyr`
us_counties <- us_counties %>%
  filter(!STATEFP %in% c("02", "15", "66", "72", "69", "78", "60"))
#check our work!
plot(us_counties)
```

Nice! It looks like we've created the desired outcome so far: a map of US counties we wish to learn more about. Now that we have an SPDF, or SpatialPolygonsDataFrame, it's time to get some data to visualize. 